---
title: "Data Case Prep - Exercises"
author: "Student XXXX"
format:
  html:
    page-layout: full
editor: visual
---

# About this template

To help you structure your submission, I have provided a Quarto mock template, which is already available for you to use. This template is designed to help you seamlessly integrate your code and analysis, ensuring a clear and organized presentation of your work. Feel free to use it as a starting point to format your responses effectively. You can include your code inside the `R` code chunks. To run a specific code chunk, select all lines that apply and hit `Ctrl+Enter` (alternatively, click on the *Run Current Chunk* button at the top-right corner of the code chunk. Alternatively, to render the Quarto altogether, ckick on the *Render* button (shortcut: `Ctrl+Shift+K`).

You can copy-paste the code chunk below each time you need to manipulate data. To provide your explanation and interpretation of the results for each question, you can write outside of the `R` coding chunk.

```{r}
#| echo: true
#| eval: true

### Enter your code here

```

Enter your analysis here.

## Tech-setup

Before you start, make sure that you have your `R` session correctly configured with all the following packages by running the code below:

```{r}
#| warning: false
#| message: false

# Package names
packages <- c("tidyverse","tidyquant","tidymodels","xts", "glue","scales", "ggthemes")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())

if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Load all packages
invisible(lapply(packages, library, character.only = TRUE))
```

Alternatively, you can simply call:

```{r}
#| eval: false #I don't want R to evaluate this code; it is just for presentation purposes
#| echo: true #I want this code chunk to display on the Quarto Document

#Install if not already available
  install.packages('tidyverse')
  install.packages('tidyquant')
  install.packages('glue')
  install.packages('scales')
  install.packages('ggthemes')

#Load
  library(tidyverse)
  library(tidyquant)
  library(tidymodels)
  library(glue)
  library(scales)
  library(ggthemes)
  
```

